Union and intersection contracts are hard, actually

### 2021-06-10 paper.tex
String -> Str, [] -> List

This same property is desirable of contracts as well, otherwise we
lose an important benefits of dynamic checking. Preferably, we want to
be able to extend the universe of contracts with user-defined
predicates (sometimes called flat contracts).

@@

Another application of intersections shows up when intersecting
functions: it can be used to encode overloading.
For instance, take a look at Figure~\ref{fig:intersection-overloading-ex}, where the function
\nickel{duplicate} works both as a function to duplicate
arrays, as well as a function to duplicate strings.
This is particularly useful when using unions, since it's a
good way to express that a function can deal with different
shapes of data.
For instance, in the same figure, \nickel{duplicate} is (correctly) called
on a value of type \nickel{[String] @| String}.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let duplicate
    | (List String -> List String)
    @& (String -> String) =
  fun x => x ++ x in
let text | [String] @| String = ... in
duplicate text
\end{lstlisting}
\caption{Duplicating an array of Strings or a String}
\label{fig:intersection-overloading-ex}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let c = contracts.frompred (fun f =>
  f 0 == 0) in
let g | (string -> string) @& c
      = fun x => x
\end{lstlisting}
\caption{intersection and user defined contracts}
\label{fig:inter-flat-contracts}
\end{figure}

Once again, intersection contracts introduce side-effects in the picture. What's
more, these side-effects interact with user-defined contracts in a non-trivial
way, while they are an important feature for validation.
\unsure{(Yann) Should we talk about thiemann here, who do support user-defined
contracts?}
% But this should not have happened, since in the program of Figure
% \ref{fig:inter-flat-contracts} \nickel{g} is never applied to \nickel{0}, in
% fact, it is never applied to anything.  Checking if \nickel{C} has already
% failed, and marking \nickel{String -> String} as failed negatively on \nickel{l}
% is not enough, the contract \nickel{String -> String} should not have been used
% inside the evaluation of \nickel{C}, which adds an extra layer of complexity:
% contracts should be guarded, and execution depends on the shape of the
% surrounding evaluation context.

@@

\begin{figure}[h]
\begin{lstlisting}[language=nickel,title={Original program}]
let C = contracts.fromPred (fun f =>
  f 0 == 0) in
let g | (String -> String) @& C
      = fun x => x
in g 0
\end{lstlisting}
\begin{lstlisting}[numbers=none,title={First component}]
g 0 | Str -> Str
\end{lstlisting}
\begin{lstlisting}[numbers=none,title={Second component}]
g 0 | C
\end{lstlisting}
\caption{Intersection and user defined contracts}
\label{fig:inter-flat-contracts-2}
\end{figure}

>>>>

This same property is desirable of contracts as well, otherwise we
lose an important benefit of dynamic checking. Preferably, we want to
be able to extend the universe of contracts with user-defined
predicates (sometimes called flat contracts).

@@

Another application of intersections shows up when intersecting
functions: it can be used to encode overloading.
For instance, take a look at Figure~\ref{fig:intersection-overloading-ex}, where the function
\nickel{duplicate} works both as a function to duplicate
arrays, as well as a function to duplicate strings.
This is particularly useful when using unions, since it's a
good way to express that a function can deal with different
shapes of data.
For instance, in the same figure, \nickel{duplicate} is (correctly) called
on a value of type \nickel{List Str @| Str}.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let duplicate
    | (List Str -> List Str)
    @& (Str -> Str) =
  fun x => x ++ x in
let text | List Str @| Str = ... in
duplicate text
\end{lstlisting}
\caption{Duplicating an array of Strings or a String}
\label{fig:intersection-overloading-ex}
\end{figure}

@@

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let c = contracts.frompred (fun f =>
  f 0 == 0) in
let g | (Str -> Str) @& c
      = fun x => x
\end{lstlisting}
\caption{intersection and user defined contracts}
\label{fig:inter-flat-contracts}
\end{figure}

Once again, intersection contracts introduce side-effects in the picture. What's
more, these side-effects interact with user-defined contracts in a non-trivial
way, while they are an important feature for validation.
\unsure{(Yann) Should we talk about thiemann here, who do support user-defined
contracts?}
% But this should not have happened, since in the program of Figure
% \ref{fig:inter-flat-contracts} \nickel{g} is never applied to \nickel{0}, in
% fact, it is never applied to anything.  Checking if \nickel{C} has already
% failed, and marking \nickel{Str -> Str} as failed negatively on \nickel{l}
% is not enough, the contract \nickel{Str -> Str} should not have been used
% inside the evaluation of \nickel{C}, which adds an extra layer of complexity:
% contracts should be guarded, and execution depends on the shape of the
% surrounding evaluation context.

@@

\begin{figure}[h]
\begin{lstlisting}[language=nickel,title={Original program}]
let C = contracts.fromPred (fun f =>
  f 0 == 0) in
let g | (Str -> Str) @& C
      = fun x => x
in g 0
\end{lstlisting}
\begin{lstlisting}[numbers=none,title={First component}]
g 0 | Str -> Str
\end{lstlisting}
\begin{lstlisting}[numbers=none,title={Second component}]
g 0 | C
\end{lstlisting}
\caption{Intersection and user defined contracts}
\label{fig:inter-flat-contracts-2}
\end{figure}
